Algebraic Geometry\vspace{-1ex}

### 2019-10-29 Algebraic Geometry/Algebraic-Geometry.tex
update

>>>>

\section{October 28th, 2019}
Max heard some things from the feedback: he is interested in seeing if he can't help us 
put things together at the end of class. That should help!

\subsection{Products}
Today we are going to be talking about products (and fiber products) in the category of varieties. 
\begin{defn}
	Let $\calC$ be a category and $X,Y\in\calC$. Then the product is the usual thing with the usual universal property.
	\begin{center}
		\begin{tikzcd}
			& X\prod Y\ar[dl,"\pi_X"]\ar[dr,"\pi_Y"] &\\
			X & & Y\\
			& A\ar[ul]\ar[ur]\ar[dashed,uu] &
		\end{tikzcd}
	\end{center}
\end{defn}
\begin{rmk}
	Notice that we can also ``think in terms of functors.'' THen we can define the product of $h_X$ and $h_Y$ as $h_X\times h_Y$ 
	in terms of the (Cartesian) product of sets.
\end{rmk}
Also recall the definition of a fiber product. Notice that if we translate back to functors, we are again trying to figure out what to 
make of the set 
\[h_X\times_{h_Z}h_Y.\]
If $f$ and $g$ are maps from $h_X$ and $h_Y$, respectively, into $h_Z$, then the fiber over $x\in h_Z$ in $h_X\times_{h_Z}h_Y$
is precisely the product of the fibers (whhaaaaat \textit{like a fibered product}) in $f$ and $g$.

\begin{prob}
	If $\calC$ has a final object $\Omega$, then $X\times Y=X\times_\Omega Y$. THink about this.
\end{prob}
\begin{thm}
	Fiber products exist in the category of varieties.
\end{thm}
Today we are going to be discussing the affine case: given $X\to Z\leftarrow Y$ of affines, constructing 
$X\times_Z Y$. To start, let $Z=\Spec k$. We are looking to study the functor $h_X\times h_Y$ and then show representability.

Let's look at this pointwise: let $X=\Spec A$ and $Y=\Spec B$. Then
\[h_X(S)\times h_Y(S)=\Hom(S,X)\times \Hom(S,Y)=\Hom(A,\O_S)\times\Hom(B,\O_S)=\Hom(A\otimes B,\O_S)=h_{\Spec A\otimes B}(S)\]
\begin{prob}
	Show that the coproduct in $\Alg_k$ is the tensor product.
\end{prob}
\begin{rmk}
	For the above problem, notice that $A\otimes B$ is odd notation because it is not just the underlying vector space, but carries with it 
	an algebra structure.
\end{rmk}
\begin{prob}
	Is the tensor product of a finitely-generated algebra finitely generated? Can you do it without coordinates?
\end{prob}

But notice for what we wrote above to be ``okay'' we need that $A\otimes_k B$ is a reduced finitely generated $k$-algebra. The finite generation 
isn't too hard to see. The reducedness comes from the following (in Kempf):

We ahve a map $A\otimes B\to \{|\Spec A|\times |\Spec B|\to k\}$ as set of functions. The claim is that this map is injective.
The idea is that if $\sum_i f_i\otimes g_i\mapsto 0$, we can assume that the $f_i$ and $g_i$ are linearly independent.

But then for all $x,$ we get $\sum_i f_i(x)g_i=0:\Spec B\to k$. But then since the $g_i$ are linearly independent,
this means the $f_i(x)$ are all zero. Thus the nullstellensatz tells us that all $f_i=0$.
